The Heavy Photon Search Test Detector

### 2013-11-03 hps-testrun-nim.tex
Changing from F to C. Small changes to text.

%Step\subsubsection{PbWO$_4$ Crystals}
The HPS test run ECal uses PbWO$_{4}$ crystals, APDs, and preamplifier boards from the inner calorimeter (IC) of the CLAS detector~\cite{2008erhm.conf..421N}.  In Fig.~\ref{fig:ecal-module} a schematic view of the module assembly is shown. Lead-tungstate crystals are $160$ mm long and are tapered with front face of $13.3\times13.3$~mm$^2$, tapered to $16\times16$~mm$^2$ at the rear face. The crystals were fabricated with very tight tolerances: $19~\mu$m for 13.3~mm dimension and $13~\mu$m for the 16~mm dimension. Each crystal is wrapped in VM2000 multilayer polymer mirror film.
The APDs (Hamamatsu S8664-55) with $5\times5$~mm$^2$ active area and $75\%$ quantum efficiency were glued on the rear face of the crystals using MeltMount 1.7 thermal plastic adhesive. The low gain of APDs ($\sim 200$) was compensated with custom made preamplifier boards, which provide factor 2000 amplification of the ADP signal.

\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/ecal-module-schematic.png}
\caption{Schematic vied of an ECal module.}
\label{fig:ecal-module}
}
\end{center}
\end{figure}

@@

During the test run ECal was mounted downstream of the analyzing dipole magnet at the distance of about $147$~cm from the 
upstream edge of the magnet. 
%The two ECal modules are positioned just above and below the ECal vacuum chamber as shown in bottom panel of Fig.~\ref{fig:ecal}. 
%The primary electron beam, scattered electrons and bremsstrahlung photons pass through the system in vacuum. 
The gap between innermost edge of the crystals of two parts was just $7.4$~cm from the beam. The temperature stability of the system during the test run was better than $1^\circ~$F, the temperature uniformity inside the box was better than  $4^\circ~$F.

@@

\subsection{Signal readout}
\label{sec:fadc}
Amplified APD signal from the motherboard is sent to a $2:1$ signal splitter. After the split, 2/3 of the signal is sent a discriminator then to a Time-to-Digital 
Converter (TDC). The 1/3 of the signal is fed to a single channel of 16-channel JLAB Flash Amplitude-to-Digital Converter VXS module (FADC250)~\cite{fadc}, see Fig.\ref{fig:fadc}.  
\begin{figure}[t]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/FADC250_Photo_001.jpg}
\caption{A Jefferson Lab FADC250 VXS module.}
\label{fig:fadc}
}
\end{center}
\end{figure}
Two 20-slot VXS crates with 14 FADC bards were needed to accommodate the system: one for each half of the ECal with 221 
channels.

The FADCs store 12-bit digitized samples at 250~MHz in 8~$\mu$s deep pipelines. 
When a trigger is received, the appropriate part of the pipeline is accessed for the readout. HPS used FADCs in the integration mode. A predefined number of samples of the signal that exceeds a predefined threshold have been summed and stored in 17-bit register for readout. The number of samples for integration was defined by two parameters that could be set separately, the number of samples integrated before the signal crossed threshold ($NSB$) and  the number of samples integrated after the signal crossed threshold ($NSA$), see Fig.\ref{fig:hps_trigger_data}. 
This scheme significantly compresses the data input 
to the FADC. During data analysis, a pedestal value is subtracted to obtain the actual summed energy.

\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/hps_trigger_data}
\caption{\small{FADC data paths}}
\label{fig:hps_trigger_data}
}
\end{center}
\end{figure}

subsection{Calibration and Performance}
\label{sec:ecal_calibration}

For various reasons only 385 out of 442 modules (87\%) were operational during the test run.  Of 442 modules, 39 were disabled or disconnected and were not read out by the DAQ. 
13 of these were not read out because of a shortage of FADC readout boards (see Sec.~\ref{sec:fadc}).
The remainder either had no HV bias on the APD, or were disabled in the FADC software due to noise.
In the data, we identified two types of abnormal channels. 
One FADC was not sending trigger signals correctly, resulting in low efficiency.
 This affected the 13 channels read out by that FADC. 5 channels were diagnosed as noisy because they 
 had a high incidence of hits out of coincidence with the trigger. A large number of channels were originally 
 misidentified as noisy because they had much higher hit occupancy than neighboring channels.
Gain calibration shows that these channels have high gain (and thus lower energy threshold) but are otherwise normal. The abnormal channels were ignored in analysis in order to simplify comparison with Monte Carlo. This leaves 385 useful channels---87\% of the ECal.

The noise level of the ECal modules were measured to be approximately 15~MeV. The ECal was operated 
with an effective readout threshold of 70~MeV.

Modules above thresholds are clustered using a simple robust algorithm described in Sec.~\ref{sec:trigger}.

%\subsubsection{Calibration}
%\label{sec:ecal_calibration}
The noise and pedestal of the readout chain are calibrated by running the ECal readout in a mode where the 
preamplifier output is sampled every 4~ns in a time window of 100 samples: by looking at a part of the 
window before the hit, we calibrate the readout channel. The The readout threshold

We calibrate gain of the individual ECal channels using the SVT measurement of track momentum and 
comparison to Monte Carlo simulation.  We disable all SVT and ECal channels in the simulation that were 
inoperable or noisy in the test run, so any efficiency or bias effects that affect the real data should be reflected 
in the simulation as well; then we use a formula to compute the ``weighted E/p'' for a crystal, representing the average E/p for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$
, and iteratively adjust the gains until the weighted E/p is equal for test run data and simulation.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/eoverp_simulation}
	\caption{Weighted E/p from Monte Carlo simulation after gain calibration.}
	\label{fig:gains}
}
\end{center}
\end{figure}

These gains can then be used to convert from ADC counts in a channel to the energy deposited into that ECal 
crystal.
The other information needed to find the energy of an incident particle is the sampling fraction---the ratio of 
energy read out from crystals to energy of an incident particle. The conventional sampling fraction, 
the fraction of incident energy that is deposited in crystals, is approximately 0.9  for the ECal and less at the 
edges. For our readout, there is additional energy lost because crystals under the readout threshold are not 
read out. The weighted E/p used in calibration (see Figure \ref{fig:gains}) is an approximate measurement of sampling fraction, but the sampling fraction is energy-dependent because of the effect of readout threshold. 
A full computation of sampling fraction can be done using simulation.

%Step\subsection{Monitoring}

@@

During the test run, for the most part trigger system run in single cluster mode and was tested to work for up to 20 kHz trigger rates.

\subsection{Data Acquisition and Online Computing}
\label{sec:daq}
For the ECal, every VXS crate contains a Readout Controller (ROC)
that collects digitized information, processes it, and sends it on to the Event Builder (EB). The ROC is a single 
blade Intel-based CPU module running DAQ software under CentOS Linux OS. For the SVT ATCA system, 
the ROC application runs on an embedded processor situated on the ATCA main board. The EB assembles 
information from the SVT and ECal ROCs into a single event which is passed to the Event 
Recorder (ER) that writes it to a RAID5-based data storage system capable of handling up to 100 MB/s. The 
EB and other critical components run on multicore Intel-based multi-CPU servers. The DAQ network system is 
a Foundry router providing high-speed connections between the DAQ components and to the JLab 
computing facility. The SVT ROC, which must handle large data volumes, has a 10 Gbit/s link to the Foundry 
router, while a 1 Gbit/s link is adequate for the ECal. A 10 Gbit/s uplink to the JLab 
computing facility is used for long-term storage.

@@

\subsubsection{Performance}
As discussed above, the trigger and DAQ integrate pulses differently to measure hit energy. The trigger 
integrates using a time-over-threshold window, and the DAQ readout integrates using a constant window 
(5 samples before and 30 samples after a threshold crossing). For every event, the trigger reports the 
trigger decision as a bit mask (top trigger, bottom trigger, or both) and the time the trigger fires.

@@

\subsection{Conclusion}
In summary, the accurate modeling of the MCS is fundamental to estimate occupancies and trigger rates 
for HPS. \egs{} predicts the correct angular distribution across all converter thicknesses while 
\geant{} overestimates the rates; with the disagreement increasing  with larger converter thickness. 
This preliminary result verifies similar studies [] and gives confidence in our modeling of the MCS 
using \egs{} for evaluating the physics reach of HPS.

>>>>

%Step\subsubsection{PbWO$_4$ Crystals}
The HPS test run ECal uses PbWO$_{4}$ crystals, APDs, and preamplifier boards from the inner calorimeter (IC) 
of the CLAS detector~\cite{2008erhm.conf..421N}.  In Fig.~\ref{fig:ecal-module} a schematic view of the module 
assembly is shown. Lead-tungstate crystals are $160$ mm long and are tapered with front face of 
$13.3\times13.3$~mm$^2$, tapered to $16\times16$~mm$^2$ at the rear face. The crystals were fabricated with 
very tight tolerances: $19~\mu$m for 13.3~mm dimension and $13~\mu$m for the 16~mm dimension. Each crystal 
is wrapped in VM2000 multilayer polymer mirror film.
The APDs (Hamamatsu S8664-55) with $5\times5$~mm$^2$ active area and $75\%$ quantum efficiency were 
glued on the rear face of the crystals using MeltMount 1.7 thermal plastic adhesive. The low gain of APDs ($\sim 
200$) was compensated with custom made preamplifier boards, which provide factor 2000 amplification of the 
ADP signal. 
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/ecal-module-schematic.png}
\caption{Schematic vied of an ECal module.}
\label{fig:ecal-module}
}
\end{center}
\end{figure}

@@

During the test run ECal was mounted downstream of the analyzing dipole magnet at the distance of about 
$147$~cm from the 
upstream edge of the magnet. 
%The two ECal modules are positioned just above and below the ECal vacuum chamber as shown in bottom panel of Fig.~\ref{fig:ecal}. 
%The primary electron beam, scattered electrons and bremsstrahlung photons pass through the system in vacuum. 
The gap between innermost edge of the crystals of two parts was just $7.4$~cm from the beam. The temperature 
stability of the system during the test run was better than $1^\circ~$F, the temperature uniformity inside the box 
was better than $3^\circ~$C.

@@

\subsection{Signal readout}
\label{sec:fadc}
Amplified APD signal from the motherboard is sent to a $2:1$ signal splitter. After the split, 2/3 of the signal is sent 
a discriminator then to a Time-to-Digital Converter. The 1/3 of the signal is fed to a single channel of 16-channel 
JLab Flash Amplitude-to-Digital Converter VXS module (FADC250)~\cite{fadc} shown in Fig.\ref{fig:fadc}.  
\begin{figure}[t]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/FADC250_Photo_001.jpg}
\caption{A JLab FADC250 VXS module.}
\label{fig:fadc}
}
\end{center}
\end{figure}
Two 20-slot VXS crates with 14 FADC bards were needed to accommodate the system: one for each half of the 
ECal with 221 channels.

The FADCs store 12-bit digitized samples at 250~MSPS in 8~$\mu$s deep pipelines. 
When a trigger is received, the appropriate part of the pipeline is accessed for the readout. HPS used FADCs in 
the co-called integration mode. A predefined number of samples of the signal that exceeds a predefined threshold 
have been summed and stored in 17-bit register for readout. The number of samples for integration was defined 
by two parameters that could be set separately, the number of samples integrated before the signal crossed 
threshold ($NSB$) and  the number of samples integrated after the signal crossed threshold ($NSA$), 
see Fig.~\ref{fig:hps_trigger_data}. 
This scheme significantly compresses the data input 
to the FADC. During data analysis, a pedestal value is subtracted to obtain the actual summed energy.
\begin{figure}[]
\begin{center}
{\small
\includegraphics[width=7cm]{figures/hps_trigger_data}
\caption{\small{FADC data paths}}
\label{fig:hps_trigger_data}
}
\end{center}
\end{figure}

\subsection{Calibration and Performance}
\label{sec:ecal_calibration}

Only 385 out of 442 modules (87\%) were operational during the test run.  Of 442 modules, 39 were disabled or 
disconnected and were not read out by the DAQ. 
13 of these were not read out because of a shortage of FADC readout boards.
The remainder either had no HV bias on the APD, or were disabled in the FADC software due to noise.
In the data, we identified two types of abnormal channels. 
One FADC was not sending trigger signals correctly, resulting in low efficiency.
 This affected the 13 channels read out by that FADC. 5 channels were diagnosed as noisy because they 
 had a high incidence of hits out of coincidence with the trigger. A large number of channels were originally 
 misidentified as noisy because they had much higher hit occupancy than neighboring channels.
Gain calibration shows that these channels have high gain (and thus lower energy threshold) but are otherwise 
normal. The abnormal channels were ignored in analysis in order to simplify comparison with Monte Carlo. This 
leaves 385 useful channels---87\% of the ECal.

The noise level of the ECal modules were measured to be approximately 15~MeV. The ECal was operated 
with an effective readout threshold of 70~MeV. Modules above thresholds are clustered using a simple robust 
algorithm described in Sec.~\ref{sec:trigger}.

%\subsubsection{Calibration}
%\label{sec:ecal_calibration}
The noise and pedestal of the readout chain are calibrated by running the ECal readout in a mode where the 
preamplifier output is sampled every 4~ns in a time window of 100 samples: by looking at a part of the 
window before the hit,  the readout channel pedestal can be obtained. 
We calibrate gain of the individual ECal channels using the SVT measurement of track momentum and 
comparison to Monte Carlo simulation.  We disable all SVT and ECal channels in the simulation that were 
inoperable or noisy in the test run, so any efficiency or bias effects that affect the real data should be reflected 
in the simulation as well; then we use a formula to compute the ``weighted E/p'' for a crystal, representing the average E/p for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$
, and iteratively adjust the gains until the weighted E/p is equal for test run data and simulation.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/eoverp_simulation}
	\caption{Weighted E/p from Monte Carlo simulation after gain calibration.}
	\label{fig:gains}
}
\end{center}
\end{figure}
These gains can then be used to convert from ADC counts in a channel to the energy deposited into that ECal 
crystal.
The other information needed to find the energy of an incident particle is the sampling fraction---the ratio of 
energy read out from crystals to energy of an incident particle. The conventional sampling fraction, 
the fraction of incident energy that is deposited in crystals, is approximately 0.9  for the center of the ECal and less 
at the 
edges. For our readout, there is additional energy lost because crystals under the readout threshold are not 
read out. The weighted E/p used in calibration (see Figure \ref{fig:gains}) is an approximate measurement of sampling fraction, but the sampling fraction is energy-dependent because of the effect of readout threshold. 
A full computation of sampling fraction can be done using simulation.

@@

During the test run, for the most part trigger system run in single cluster mode and was tested to work for up to 
20~kHz trigger rates.

\subsection{Data Acquisition and Online Computing}
\label{sec:daq}
For the ECal, every VXS crate contains a Readout Controller (ROC)
that collects digitized information, processes it, and sends it on to the Event Builder (EB). The ROC is a single 
blade Intel-based CPU module running DAQ software under CentOS Linux OS. For the SVT DAQ system, 
the ROC application runs on a dedicated DAQ desktop computer. The EB assembles 
information from the SVT and Cal ROCs into a single event which is passed to the Event 
Recorder (ER) that writes it to a RAID5-based data storage system capable of handling up to 100~MB/s. The 
EB and other critical components run on multicore Intel-based multi-CPU servers. The DAQ network system is 
a Foundry router providing high-speed connections between the DAQ components and to the JLab 
computing facility. The SVT ROC, which must handle large data volumes, has a 10~Gb/s link to the Foundry 
router, while a 1~Gb/s link is adequate for the ECal. A 10~Gb/s uplink to the JLab 
computing facility is used for long-term storage.

@@

\subsection{Performance}
As discussed above, the trigger and DAQ integrate pulses differently to measure hit energy. The trigger 
integrates using a time-over-threshold window, and the DAQ readout integrates using a constant window 
(5 samples before and 30 samples after a threshold crossing). For every event, the trigger reports the 
trigger decision as a bit mask (top trigger, bottom trigger, or both) and the time the trigger fires.

@@

\subsection{Conclusion}
In summary, the accurate modeling of the MCS is fundamental to estimate occupancies and trigger rates 
for HPS. \egs{} predicts the correct angular distribution across all converter thicknesses as expected while 
\geant{}, using the so-called Urban model for multiple scattering overestimates the rates; with the disagreement 
increasing  with larger converter thickness. This preliminary result verifies similar studies [] and gives confidence 
in our modeling of the MCS 
using \egs{} for evaluating the physics reach of HPS.
